Redshift Evolution of the Underlying Type Ia Supernova Stretch Distribution

### 2020-10-14 Nicolas.tex
Integrated YC comments

\usepackage{txfonts}
\usepackage{amsmath, mathtools}
\usepackage[colorlinks,breaklinks]{hyperref}
\hypersetup{linkcolor=blue,citecolor=blue,filecolor=black,urlcolor=blue}

@@

\abstract{The true nature of type Ia supernovae (SNe~Ia) remains largely
    unknown, and as survey statistics increase, the question of astrophysical
    systematic uncertainties rises, notably that of the SN~Ia population
    evolution. In this paper, we study the dependence with redshift of the SN~Ia
    \texttt{SALT2.4} lightcurve stretch, a purely intrinsic SN property, to
    probe its potential redshift drift. The SN stretch has been shown to
    strongly correlate with the SN environment, notably with stellar age
    tracers. We model the underlying stretch distribution as a function of
    redshift, using the evolution of the fraction of young and old SNe~Ia as
    predicted by \cite{rigault2018}, and assuming constant underlying stretch
    distribution for each age population made of Gaussian mixtures. We test our
    prediction against published samples chosen to have negligible magnitude
    selection effects, so that any observed change is indeed of astrophysical
    and not observational origin. We clearly demonstrate that the underlying
    SN~Ia stretch distribution is evolving as a function of redshift, and that
    the young/old drifting model is a much better description of the data than
    any time-constant model, including the sample-based asymmetric distributions
    usually used to correct Malmquist bias. The favored underlying stretch model
    is the bimodal one derived from \cite{rigault2018}: a high-stretch mode
    shared by both young and old environments, and a low-stretch mode exclusive
    to old environments. The precise impact of the redshift evolution of the SN
    Ia population intrinsic properties on cosmology remains to be studied. Yet,
    the astrophysical drift of the SN stretch distribution does affect current
    Malmquist bias corrections and thereby distances derived from SN affected by
    selection effects. We highlight that such a bias will increase with surveys
    covering increasingly larger redshift ranges, which is particularly
    important for LSST.}

@@

\section{Complete Sample Construction}\label{sec:sample}
\subsection{\nn{Applying redshift cuts}}\label{ssec:cuts}

@@

SNLS typically acquired SNe~Ia in the redshift range $0.4<z<0.8$; at these
redshifts, the rest-frame Bessel $B$ band roughly corresponds to the SNLS $i$
filter, that has a $5\sigma$ depth of
24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
final release website.}}. This converts to a $z_{\lim}=0.60$, in agreement with
\cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. Fig.~14 of
\cite[][see their Section~5]{perrett2010} suggests a lower limit of
$z_{\lim}=0.55$; both limits will be considered, as discussed below.

Similarly, PS1 observed SNe~Ia in the range $0.2<z<0.4$, their $g$-band
$5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields to $z_{\lim}=0.30$ in
agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. This figure is also
suggestive of a more conservative $z_{\lim}=0.27$.

In a similar redshift range, SDSS has a limiting magnitude of 22.5
\citep{dilday2008, sako2008}, which would lead to a $z_{\lim}=0.24$. However,
the SDSS surveys were more sensitive to limited spectroscopic resources;
\cite[][see their Section~2]{kessler2009} pointed out that during the first year
of SDSS, SNe~Ia with $r<20.5$~mag were favored for spectroscopic follow-up,
corresponding to a redshift cut at $0.15$. For the rest of the SDSS survey,
additional spectroscopic resources were available, and \cite{kessler2009} and
\cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$. Following
these analyses, we will use $z_{\lim}=0.2$ as the baseline SDSS redshift limit.

The sample selection is summarized in Table~\ref{tab:sample}, and the redshift
distribution of these three surveys is shown in Fig.~\ref{fig:cuts}. As
expected, the selected redshift limits roughly correspond to the peak of these
histograms. We show in Section~\ref{sec:results} that this sample selection
indeed provides a subset of SNe~Ia with insignificant selection effects when
compared to state-of-the-art Malmquist correction techniques.

\begin{table}
    \centering
    \caption{Composition of the SNe~Ia dataset used in this analysis.
    Conservative cuts are indicated in parentheses.}
    \label{tab:sample}
    \begin{tabular}{l c c}
        \hline\hline
        Survey & $z_{\lim}$ & $N_{\mathrm{SN}}$ \\
        \hline
        SNf   & --          & 114 \\
        SDSS  & 0.20 (0.15) & 167 (82)\\
        PS1   & 0.30 (0.27) & 160 (122)\\
        SNLS  & 0.60 (0.55) & 102 (78)\\
        HST   & --          & 26 \\
        \hline
        Total & --          & 569 (422) \\
        \hline
    \end{tabular}
\end{table}

@@

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory,][]{aldering2002} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
and~4.2.2; SNe~Ia time series are published in \citealt{saunders2020}, see also
\citealt{aldering2020}). The spectroscopic follow-up was done over a redshift
range of $0.02 < z < 0.09$~\citep[as in][]{rigault2018}, while the search was
much deeper. These 114 SNfactory SNe~Ia are thus in the volume limited part of
the survey (Aldering et al., in prep.), and are therefore assumed to be a random
sampling of the underlying SN population. The SNfactory sample is particularly
useful for studying SN property drift, as it enables us to have a large SN~Ia
sample at $z<0.1$.

@@

We present the stretch distribution and redshift histogram of these five surveys
up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. The evolution of the
mean stretch is also shown in Fig.~\ref{fig:modelall}, where the data are split
in redshift bins of regular sample size. We see that SNe~Ia at higher redshift
have on average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at
lower redshift ($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying
stretch distribution is drifting.

@@

\subsection{\mr{Testing the construction of a volume limited sample}}
\label{ssec:verify}

\mr{In this analysis, we have built a volume limited sample from a set of
    magnitude limited ones, using a simple redshift limit. An ongoing follow up
    analysis will make use of the SNANA package \citep{SNANA} to account for
    more subtle selection functions. However, the strength of the current
    approach is that it is independent of potential inaccuracy in the modeling
    of the complex selection functions of the aforementioned dataset. Yet, it is
    necessary to test if the sample we have created could show sign of leftover
sample selections.}

\mr{To do so, we compare the stretch and color distributions of the SNe~Ia in
    our sample between datasets of overlapping redshifts. If indeed we have
    built a volume limited sample, the distributions of any SNe~Ia parameter at
    a given redshift should not depend of their origin surveys. When considering
    a redshift range, this has to be narrow enough such that any drift would be
    negligible. The two samples that overlap the most in redshift are PS1 and
    SDSS in the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}).
    This corresponds to the 146 SNe~Ia from SDSS at the highest redshift end and
    thus potentially the most affected by potential selection effect leftover
    (see the corresponding discussion in section~\ref{sec:sample}). In that same
    redshift range, PS1 has 52 SNe~Ia, which are at in the lowest redshift bin
    and thus unlikely to have any selection issue. The Fig.~\ref{fig:distrib}
    present the stretch and color distribution of both, together with p-values
    of Kolmogorov-Smirnov tests that test if the $x_1$ and $c$ of the SDSS and
    PS1 samples at $0.10<z<0.20$ arise from the same underlying distribution.
    Those p-values are high ($>10\%$), and if one cannot conclude from this test
    that they SNe~Ia are from the same underlying distribution, one has no
    ground to claim they are not. This is in agreement with the visual
    impression from Fig.  ~\ref{fig:distrib}.  A similar analysis is doable for
    PS1 and SNLS in their redshift range $0.20 < z < 0.30$ where the same
    conclusion could be drawn, i.e., that there is no sign of discrepancy
    between in the stretch and color distributions between the low- and high-end
    of the SNLS and PS1 datasets, respectively. Nonetheless, the small size of
    the SNLS dataset at $z < 0.30$ (12 SNe~Ia vs. 90 for PS1) limits the
    interest of the test as only strong deviation would be noticeable. Extending
the redshift range to $0.20 < z < 0.40$ (though we have no PS1 data above 0.3)
enables to get 31 SNe~Ia from SNLS in total, the stretch and color p-values
remains high, at 34\% and 4\% respectively.}

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/both-cut_SDSS_SNLS_PS1.pdf}
    \caption{\mr{$x_1$ and $c$ distribution histograms of different surveys
    overlapping in redshift. \emph{Facing up} SDSS and PS1 within the redshift
    range $0.10 < z < 0.20$ ; \emph{facing down} PS1 and SNLS within the
    redshift range $0.20<z<0.3$. Error bars show the Poisson noise. Our stretch
    "Base" modeling is illustrated in orange at the mean redshift of the
    redshift ranges, $0.15$ and $0.25$, respectively. P-values of the
    Kolmogorov-Smirnov tests are indicated on the top (bottom) of each panel
    showing no sign that the SDSS and PS1 (PS1 and SNLS) $x_1$ and $c$
    distributions are not drawn from the same underlying distributions.}}
    \label{fig:distrib}
\end{figure}

\mr{We highlight that the SNe~Ia color is more prone to selection effects than
    stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g., Fig.~3 of
    \cite{kessler2017}. Hence, since the color shows no sign of leftover
    selection effects, it further supports our claim that our simple
    redshift-based selection criteria is sufficient to test the redshift drift
of the SN~Ia stretch distribution.}

@@

Results of fitting this model to all the 569 (resp. 422) SNe from the fiducial
(resp. conservative) sample are given Table~\ref{tab:modelresults}, and the
predicted redshift evolution of mean stretch (expected $x_1$ given the
distribution of Eq.~\ref{eq:stretchz}) illustrated as a blue band in
Fig.~\ref{fig:modelall} accounting for parameters errors and their covariances.
We see in this figure that the measured mean SN~Ia stretch per redshift bins of
equal sample size closely follows our redshift drift modeling. This is indeed
what is expected if old environments favor low SN stretches
\citep[e.g.][]{howell2007} \emph{and} if the fraction of old SNe~Ia declines as
a function of redshift. See Section~\ref{sec:results} for a more quantitative
discussion.

@@

We also consider another intrinsically non-drifting model, \mr{the functional
    form} developed for Beams with Bias Correction \cite[BBC,][]{scolnic2016,
    kessler2017}, used in recent SN cosmological analyses
    \cite[e.g.][]{scolnic2018a, descosmopaper2019, riess2016, riess2019} to
    account for Malmquist biases. The BBC formalism assumes sample-based (hence
    intrinsically non-drifting) asymmetric Gaussian stretch distributions:
    $\mathcal{N}\left(\mu,
    \sigma_-{}^2\;\text{if}\;x_1<\mu,\;\text{else}\;\sigma_+{}^2\right)$. The
    idea behind this sample-based approach is twofold: (1) Malmquist biases are
    driven by survey properties and (2) since current surveys cover limited
    redshift ranges, having a sample-based approach covers some potential
    redshift evolution information \citep{scolnic2016, scolnic2018a}. See
    further discussion concerning BBC in Section~\ref{sec:discussion}.

@@

We report in Table~\ref{tab:bbc} the samples' $\mu_0$ and $\sigma_{\pm}$
adjusted on the nominally selection-free samples using our fiducial cuts (see
Section~\ref{sec:sample}). We find our results in close agreement with
\cite{scolnic2016} for SNLS and SDSS and with \cite{scolnic2018a} for PS1, who
derived these model parameters using the full BBC formalism, using numerous
simulations to model the selection effects (see details e.g., Section~3 of
\citealt{kessler2017}). The agreement between our fit of the asymmetric
Gaussians on the supposedly selection-free part of the samples and the results
derived using the BBC formalism supports our approach to get a sample with
negligible selection effects. If we were to use \cite{scolnic2016} and
\cite{scolnic2018a} best fit values of the $\mu_0, \sigma_{\pm}$ asymmetric
parameters for SNLS, SDSS and PS1, respectively, the $\Delta$AIC between our
Base drifting model and the BBC modeling would go even deeper from $-32$ to
$-47$. We further discuss the consequence of this result for cosmology in
Section~\ref{sec:discussion}.

@@

Our conclusions are the following:
\begin{enumerate}
    \item The underlying SN~Ia stretch distribution is significantly redshift
        dependent, as previously suggested by e.g.~\cite{howell2007}. This
        result is largely independent of details on each age-population
        modeling.

@@

In this paper, we considered a simple Gaussian mixture modeling, but additional
data free from significant Malmquist bias would enable us to refine it as
necessary. We note that samples at the low- and high-redshift ends of the Hubble
diagram would be particularly helpful for this drifting analysis; fortunately
this will soon be provided by the Zwicky Transient Facility
\citep[low-$z$,][]{bellm2019, graham2019}, and Subaru and SeeChange SNe~Ia
programs (high-$z$), respectively.

\begin{acknowledgements}
    This project has received funding from the European Research Council (ERC)
    under the European Union's Horizon 2020 research and innovation programme
    (grant agreement n 759194 - USNAC).
    This work was supported in part by the Director, Office of Science, Office
    of High Energy Physics of the U.S. Department of Energy under Contract No.
    DE-AC025CH11231.
    This project is partly financially supported by Region Rhone-Alpes-Auvergne
\end{acknowledgements}

@@

% X
% Y
% Z
\end{thebibliography}
\end{document}
end{document}

>>>>

\usepackage{txfonts}
\usepackage{ulem}
\usepackage{amsmath, mathtools}
\usepackage[colorlinks,breaklinks]{hyperref}
\hypersetup{linkcolor=blue,citecolor=blue,filecolor=black,urlcolor=blue}

@@

\abstract{The true nature of type Ia supernovae (SNe~Ia) remains largely
    unknown, and as survey statistics increase, the question of astrophysical
    systematic uncertainties rises, notably that of the SN~Ia population
    evolution. In this paper, we study the dependence with redshift of the SN~Ia
    \texttt{SALT2.4} lightcurve stretch, a purely intrinsic SN property, to
    probe its potential redshift drift. The SN stretch has been shown to
    strongly correlate with the SN environment, notably with stellar age
    tracers. We model the underlying stretch distribution as a function of
    redshift, using the evolution of the fraction of young and old SNe~Ia as
    predicted by \cite{rigault2018}, and assuming constant underlying stretch
    distribution for each age population made of Gaussian mixtures. We test our
    prediction against published samples \nn{\sout{chosen} cut} to have
    negligible magnitude selection effects, so that any observed change is
    indeed of astrophysical and not observational origin. \nn{An ongoing
    analysis with the SNANA package will test this assumption in order to refine
    the study.} We \nn{\sout{clearly}} demonstrate that the underlying SN~Ia stretch
    distribution is evolving as a function of redshift, and that the young/old
    drifting model is a much better description of the data than any time-constant
    model, including the sample-based asymmetric distributions usually used to
    correct Malmquist bias. The favored underlying stretch model is the bimodal one
    derived from \cite{rigault2018}: a high-stretch mode shared by both young and
    old environments, and a low-stretch mode exclusive to old environments. The
    precise impact of the redshift evolution of the SN Ia population intrinsic
    properties on cosmology remains to be studied. Yet, the astrophysical drift of
    the SN stretch distribution does affect current Malmquist bias corrections and
    thereby distances derived from SN affected by selection effects. We highlight
    that such a bias will increase with surveys covering increasingly larger
    redshift ranges, which is particularly important for LSST.}

@@

\section{Complete Sample Construction}\label{sec:sample}
\subsection{Applying redshift cuts}\label{ssec:cuts}

@@

SNLS typically acquired SNe~Ia in the redshift range $0.4<z<0.8$; at these
redshifts, the rest-frame Bessel $B$ band roughly corresponds to the SNLS $i$
filter, that has a $5\sigma$ depth of
24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
final release website.}}. This converts to a $z_{\lim}=0.60$, in agreement
with \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. \nn{However,
because this simple approach might easily not be enough to alleviate all
selection effects, we also take into account another, more restrictive,
redshift limit:} Fig.~14 of \cite[][see their Section~5]{perrett2010} suggests a
lower limit of $z_{\lim}=0.55$; both limits will be considered, as discussed
below.

Similarly, PS1 observed SNe~Ia in the range $0.2<z<0.4$, their $g$-band
$5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields to $z_{\lim}=0.31$ in
agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. This figure is also
suggestive of a more conservative $z_{\lim}=0.27$, \nn{similarly considered in
the following analysis as a first step to evaluate the robustness of our
sample construction.}

In a similar redshift range, SDSS has a limiting magnitude of 22.5
\citep{dilday2008, sako2008}, which would lead to a $z_{\lim}=0.24$. However,
the SDSS surveys were more sensitive to limited spectroscopic resources;
\cite[][see their Section~2]{kessler2009} pointed out that during the first year
of SDSS, SNe~Ia with $r<20.5$~mag were favored for spectroscopic follow-up,
corresponding to a redshift cut at $0.15$. For the rest of the SDSS survey,
additional spectroscopic resources were available, and \cite{kessler2009} and
\cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$. Following
these analyses, we will use $z_{\lim}=0.20$ as the baseline SDSS redshift limit,
\nn{and the first 0.15 cut to compare the results of our modeling on a more
restrictive sample.}

The sample selection is summarized in Table~\ref{tab:sample}, and the redshift
distribution of these three surveys is shown in Fig.~\ref{fig:cuts}. As
expected, the selected redshift limits roughly correspond to the peak of these
histograms. \nn{We test in Section~\ref{ssec:verify} whether we can consider
this sample to be free of selection effects. The results exposed} in
Section~\ref{sec:results} further supports that this sample selection indeed
provides a subset of SNe~Ia with insignificant selection effects when
compared to state-of-the-art Malmquist correction techniques.

\begin{table}
    \centering
    \caption{Composition of the SNe~Ia dataset used in this analysis.
    Conservative cuts are indicated in parentheses.}
    \label{tab:sample}
    \begin{tabular}{l c c}
        \hline\hline
        Survey & $z_{\lim}$ & $N_{\mathrm{SN}}$ \\
        \hline
        SNf & -- & 114 \\
        SDSS & 0.20 (0.15) & 167 (82)\\
        PS1 & 0.31 (0.27) & 160 (122)\\
        SNLS & 0.60 (0.55) & 102 (78)\\
        HST & -- & 26 \\
        \hline
        Total & -- & 569 (422) \\
        \hline
    \end{tabular}
\end{table}

@@

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory,][]{aldering2002} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
and~4.2.2; SNe~Ia time series are published in \citealt{saunders2020}, see also
\citealt{aldering2020}). The spectroscopic follow-up was done over a redshift
range of $0.02 < z < 0.09$~\citep[as in][]{rigault2018}, while the search was
much deeper. These 114 SNfactory SNe~Ia \nn{(80\% of the 141 SNe presented in
the previously-cited paper)} are thus in the volume-limited part of the
survey (Aldering et al., in prep.), and are therefore assumed to be a random
sampling of the underlying SN population. The SNfactory sample is
particularly useful for studying SN property drift, as it enables us to have
a large SN~Ia sample at $z<0.1$.

@@

We present the stretch distribution and redshift histogram of these five surveys
up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}, \nn{where the
disappearance of the low-stretch population (typically $x_1 < -1$) is what we
test in this paper}. The evolution of the mean stretch is also shown in
Fig.~\ref{fig:modelall}, where the data are split in redshift bins of regular
sample size. We see that SNe~Ia at higher redshift have on average larger
stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower redshift ($-0.17\pm
0.10$ at $z\sim0.05$), suggesting that the underlying stretch distribution is
drifting.

@@

\subsection{Testing the construction of a volume-limited sample}\label{ssec:verify}

In this analysis, we have built a volume-limited sample from a set of magnitude
limited ones, using a simple redshift limit. \nn{Even though it has the strength
to be independent of potential inaccuracy in the modeling of the complex
selection functions of the aforementioned dataset, it is suboptimal.
That fact lead to the use of conservative cuts as first indicators that there is
a drift that can't be explained by selection effects alone and a}n ongoing
follow up analysis will make use of the SNANA package \citep{SNANA} to account
for more subtle selection functions; \nn{however}, it is still necessary to test
if the sample we have created could show sign of leftover sample selection.

To do so, we compare the stretch and color distributions of the SNe~Ia in our
sample between datasets of overlapping redshifts. If indeed we have built a
volume-limited sample, the distributions of any SNe~Ia parameter at a given
redshift should not depend of their survey or origin. When considering a
redshift range, this has to be narrow enough such that any drift would be
negligible. The two samples that overlap the most in redshift are PS1 and SDSS
in the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This
corresponds to the 146 SNe~Ia from SDSS at the highest redshift end and thus the
most affected by potential selection effect leftover (see the corresponding
discussion in section~\ref{sec:sample}). In that same redshift range, PS1 has 52
SNe~Ia, which are at in the lowest redshift bin and thus unlikely to have any
selection issue. The Fig.~\ref{fig:distrib} present the stretch and color
distribution of both, together with p-values of Kolmogorov-Smirnov tests that
\nn{probe whether} the $x_1$ and $c$ of the SDSS and PS1 samples at $0.10 < z <
0.20$ arise from the same underlying distribution. Those p-values are high
($>10\%$), and if one cannot conclude from this test that these SNe~Ia are from
the same underlying distribution, one has no ground to claim they are not. This
is in agreement with the visual impression from Fig.~\ref{fig:distrib}. A
similar analysis \nn{was done} for PS1 and SNLS in the redshift range $0.20 < z
< 0.31$ where the same conclusion \nn{can} be drawn, i.e., there is no sign of
discrepancy in the stretch and color distributions between the low- and high-end
of the SNLS and PS1 datasets, respectively. Nonetheless, the small size of the
SNLS dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1) limits the interest of the
test as only strong deviation would be noticeable. Extending the redshift range
to $0.20 < z < 0.40$ (though we have no PS1 data above 0.3) enables to get 31
SNe~Ia from SNLS in total, the stretch and color p-values remains high, at 34\%
and 4\% respectively.

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/both-cut_SDSS_SNLS_PS1.pdf}
    \caption{$x_1$ and $c$ distribution histograms of different surveys
        overlapping in redshift. \textit{Facing up}: SDSS and PS1 within the
        redshift range $0.10 < z < 0.20$; \textit{facing down}: PS1 and SNLS
        within the redshift range $0.20 < z < 0.31$. Error bars show the Poisson
        noise. Our stretch ``Base'' modeling is illustrated in orange at the mean
        redshift of the redshift ranges, $0.15$ and $0.25$, respectively.
        P-values of the Kolmogorov-Smirnov tests are indicated on the top
        (bottom) of each panel showing no sign that the SDSS and PS1 (PS1 and
        SNLS) $x_1$ and $c$ distributions are not drawn from the same underlying
        distributions.}
    \label{fig:distrib}
\end{figure}

We highlight that the SNe~Ia color is more prone to selection effects than
stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g., Fig.~3 of
\cite{kessler2017}. Hence, since the color shows no sign of leftover selection
effects, it further supports our claim that our simple redshift-based selection
criteria is sufficient to test the \nn{existence of the} redshift drift of the
SN~Ia stretch distribution.

@@

Results of fitting this model to all the 569 (resp. 422) SNe from the fiducial
(resp. conservative) sample are given Table~\ref{tab:modelresults}, and the
predicted redshift evolution of mean stretch (expected $x_1$ given the
distribution of Eq.~\ref{eq:stretchz}) illustrated as a blue band in
Fig.~\ref{fig:modelall} accounting for parameters errors and their covariances.
We see in this figure that the measured mean SN~Ia stretch per redshift bins of
equal sample size closely follows our redshift drift modeling. This is indeed
what is expected if old environments favor low SN stretches
\citep[e.g.][]{howell2007} \textit{and} if the fraction of old SNe~Ia declines as
a function of redshift. See Section~\ref{sec:results} for a more quantitative
discussion.

@@

We also consider another intrinsically non-drifting model, the functional form
developed for Beams with Bias Correction \cite[BBC,][]{scolnic2016,
kessler2017}, used in recent SN cosmological analyses
\cite[e.g.][]{scolnic2018a, descosmopaper2019, riess2016, riess2019} to account
for Malmquist biases. The BBC formalism assumes sample-based (hence
intrinsically non-drifting) asymmetric Gaussian stretch distributions:
$\mathcal{N}\left(\mu, \sigma_-{}^2\; \text{if} \;x_1<\mu,\; \text{else}
\;\sigma_+{}^2\right)$. The idea behind this sample-based approach is twofold:
(1) Malmquist biases are driven by survey properties and (2) since current
surveys cover limited redshift ranges, having a sample-based approach covers
some potential redshift evolution information \citep{scolnic2016, scolnic2018a}.
See further discussion concerning BBC in Section~\ref{sec:discussion}.

@@

We report in Table~\ref{tab:bbc} \nn{our determination of} the samples' $\mu_0$
and $\sigma_{\pm}$ adjusted on the nominally selection-free samples using our
fiducial cuts (see Section~\ref{sec:sample}) \nn{when implementing an asymmetric
Gaussian fit}. We find our results in close agreement with \cite{scolnic2016}
for SNLS and SDSS and with \cite{scolnic2018a} for PS1, who derived these model
parameters using the full BBC formalism, using numerous simulations to model the
selection effects (see details e.g., Section~3 of \citealt{kessler2017}). The
agreement between our fit of the asymmetric Gaussians on the supposedly
selection-free part of the samples and the results derived using the BBC
formalism supports our approach to get a sample with negligible selection
effects. If we were to use \cite{scolnic2016} and \cite{scolnic2018a} best fit
values of the $\mu_0, \sigma_{\pm}$ asymmetric parameters for SNLS, SDSS and
PS1, respectively, the $\Delta$AIC between our Base drifting model and the BBC
modeling would go even deeper from $-32$ to $-47$. We further discuss the
consequence of this result for cosmology in Section~\ref{sec:discussion}.

@@

Our conclusions are the following:
\begin{enumerate}
    \item The underlying SN~Ia stretch distribution is significantly redshift
        dependent, as previously suggested by e.g.~\cite{howell2007} \nn{in a
        way that selection effects alone cannot explain}. This result is largely
        independent of details on each age-population modeling.

@@

In this paper, we considered a simple Gaussian mixture modeling, but additional
data free from significant Malmquist bias would enable us to refine it as
necessary. We note that samples at the low- and high-redshift ends of the Hubble
diagram would be particularly helpful for this drifting analysis; fortunately
this will soon be provided by the Zwicky Transient Facility
\citep[low-$z$,][]{bellm2019, graham2019}, and Subaru and SeeChange SNe~Ia
programs (high-$z$), respectively. \nn{Moreover, a more detailed analysis using
the SNANA framework is currently in progress in order to account for more subtle
selection functions and refine the dataset.}

\begin{acknowledgements}
    This project has received funding from the European Research Council (ERC)
    under the European Union's Horizon 2020 research and innovation programme
    (grant agreement n 759194 - USNAC).
    This work was supported in part by the Director, Office of Science, Office
    of High Energy Physics of the U.S. Department of Energy under Contract No.
    DE-AC025CH11231.
    This project is partly financially supported by Région Rhône-Alpes-Auvergne.
\end{acknowledgements}

@@

% X
% Y
% Z
\end{thebibliography}
\end{document}
